Predicting and Preventing Employee Turnover through HR Analytics

### 2024-06-19 code/eda-code/HReda.qmd
Project part 3 - First Update

# HR Data Exploration Through Tables

@@

```{r}
summary_df = skimr::skim(myhrdata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "HRsummarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

Gender Table

```{r}
#summary_df = skimr::skim(myhrdata)
#print(summary_df)

summary_df = myhrdata %>%
  count(Gender) %>%
  mutate(`Total Count` = n,
         `% of Total` = (n / sum(n)) * 100) %>%
  select(Gender, `Total Count`, `% of Total`)

@@

Tenure Distribution

```{r}
p1 = myhrdata %>%
  ggplot(aes(x = YearsAtCompany)) +
  geom_histogram(binwidth = 1, boundary = 0.5, fill = "gray", color = "black") +
  labs(x = "Tenure", y = "Count") + 
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))

@@

```{r}
p2 = myhrdata %>%
  ggplot(aes(x = TenureCategory)) +
  geom_bar(fill = "gray", color = "black") +
  labs(x = "Tenure", y = "Count", title = "Distribution of Tenure Categories") + 
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)

@@

```{r}
education_levels <- c("Below College", "College", "Bachelor's", "Master's", "Doctorate")

# Convert Education to factor with defined levels
myhrdata$Education <- factor(myhrdata$Education, levels = education_levels)

# Boxplot of MonthlyIncome by Education
p3 <- ggplot(myhrdata, aes(x = Education, y = MonthlyIncome)) +
  geom_boxplot(fill = "gray", color = "black") +
  labs(x = "Education", y = "Monthly Income", title = "Boxplot of Monthly Income by Education") + 
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1))

# Print the plot
print(p3)

@@

>>>>

## HR Data Exploration Through Tables

@@

## Data Summary

Basic summary of data.
```{r}
summary_df = skimr::skim(myhrdata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "HRsummarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

## Gender Table

```{r}

summary_df = myhrdata %>%
  count(Gender) %>%
  mutate(`Total Count` = n,
         `% of Total` = (n / sum(n)) * 100) %>% #creating new column for percentage of total
  select(Gender, `Total Count`, `% of Total`) #selecting only 3 variables

@@

## Tenure Distribution

Visualizing tenure to find distribution.
```{r}
p1 = myhrdata %>%
  ggplot(aes(x = YearsAtCompany)) +
  geom_histogram(binwidth = 1, boundary = 0.5, fill = "gray", color = "black") +
  labs(x = "Tenure", y = "Count") + 
  theme_minimal()

@@

## Distribution of Tenure

```{r}
p2 = myhrdata %>%
  ggplot(aes(x = TenureCategory)) +
  geom_bar(fill = "gray", color = "black") +
  labs(x = "Tenure", y = "Count", title = "Distribution of Tenure Categories") + 
  theme_minimal()

@@

## Education vs Monthly Income

```{r}
p3 = ggplot(myhrdata, aes(x = Education, y = MonthlyIncome)) +
  geom_boxplot(fill = "gray", color = "black") +
  labs(x = "Education", y = "Monthly Income", title = "Boxplot of Monthly Income by Education") + 
  theme_minimal()

print(p3)

@@

## Age vs Attrition

```{r}
p5 = ggplot(myhrdata, aes(x = Age, fill = Attrition)) +
  geom_histogram(binwidth = 1, position = "dodge") +
  scale_fill_manual(values = c("No" = "grey", "Yes" = "#595959")) + #selecting bar colors
  labs(title = "Age vs Attrition", x = "Age", y = "Count") +
  theme_minimal()

print(p5)

figure_file = here("results", "figures","age-attrition.png")
ggsave(filename = figure_file, plot=p5)
```

```{r}
p4 = ggplot(myhrdata, aes(x = Department, fill = Attrition)) +
  geom_bar(position = "dodge", color = "black") +
  scale_fill_manual(values = c("No" = "grey", "Yes" = "#595959")) + #selecting bar colors
  labs(title = "Department vs Attrition", x = "Department", y = "Count") +
  theme_minimal()

print(p4)

figure_file = here("results", "figures","department-attrition.png")
ggsave(filename = figure_file, plot=p4)
```

### 2024-06-19 code/processing-code/HRprocessingfile.qmd
Project part 3 - First Update

---
title: "Cleaning HR Data for Analysis"
author: "Guillermo Gallardo"
date: "Summer 2024"
output: html_document

@@

**NEED TO UPDATE THE EDUCATION FIELD**

@@

```{r}
HRprocesseddata = d3
```

>>>>

---
title: "Cleaning HR Data for Analysis"
author: "Guillermo Gallardo"
date: "Summer 2024"
output: html_document
---

@@

@@

Saving updated dataset into HRprocesseddata
```{r}
HRprocesseddata = d3
```

### 2024-06-19 products/manuscript/Manuscript.qmd
Project part 3 - First Update

#hr_data = read.csv("C:/Users/crew_/OneDrive/Escritorio/Guille/UTSA Masters/MSDA/DA6833/PracticumII-Course/guillermogallardo-final-project/data/raw-data/IBM HR Analytics Data.csv", header = TRUE) #not synced with OneDrive but my computer added that folder when I entered my email when I bought the computer
```

@@

@@

>>>>

```

@@

### **Age vs Attrition**

In the bar chart, it is evident that attrition peaks among employees aged between 28 and 32. This trend suggests that individuals in this age group may be more likely to leave the company. Possible reasons for increased turnover in this demographic could include career advancement opportunities elsewhere, desire for higher compensation, or life changes such as starting a family.

```{r}
#| label: fig-Age-Attrition
#| fig-cap: "Attrition by Employee Age"
#| echo: FALSE
knitr::include_graphics(here("results","figures","age-attrition.png"))
```

@@

```{r}

```
